---
title: "Bonus1"
author: "Uzo J Iwuagwu"
date: "June 5, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r packages}

library(ggplot2)
library(plyr)
library(vcd)
library(ggthemes)
library(caret)
#install.packages(GoodmanKruskal)
library(ResourceSelection)
library(randomForest)
library(e1071)
library(nnet)
```
Use Random Forest model type to predict wether a person makes over 50K a year

Clean and read the Adult data set

```{r Analyzing Data Set}
adult_data =read.table("C:/Users/PCL/Documents/FlowTap/FlowTap_Quizz/adult.data.txt", header = TRUE, sep = ",", col.names = c("Age", "workclass", "fnlwgt", "Education", "Educationnum", "Maritalstatus", "Occupation", "Relationship", "race", "Sex", "Capitalgain", "Capitalloss", "hoursperweek", "NativeCountry", "Class"), fill = FALSE, na.strings = "?", strip.white = TRUE)
adult_data

```



Check for missing value:
```{r Missing Values}
   any(is.na(adult_data))
```
Remove Missing Values
```{r Remove missing value}
    adult_data <- na.omit(adult_data)
      adult_data
```
Clean and Read Test data set
```{r Analyzing Test Data Set}
adult_train <- adult_data[1:15081,]
adult_test <- adult_data[15082:30161,]

```


Remember we removed the Educationnum from the logistic model due to collinearity, therefore we will remove it here

```{r Covariates}
set.seed(1234)

covariates = paste("Age", "workclass", "Education",
                    "Maritalstatus", "Occupation", "Relationship",
                    "race", "Sex", "Capitalgain", "Capitalloss","hoursperweek", "NativeCountry",
                    sep = "+")

form <- as.formula(paste("Class ~", covariates))

start_time <- proc.time()
ModRF <- randomForest(formula = form,
                      data= adult_train)
time.rf <- proc.time() - start_time
time.rf
```


```{r Out of bag Error}
ModRF
```
We see that the "out-of-bag" error, or OOB, is 18.33%:

##Test the accuracy on the training and test datasets

```{r testing Accuracy}
mean(predict(ModRF, newdata = adult_train) == adult_train$Class)
mean(predict(ModRF, newdata = adult_test) == adult_test$Class)
```
Above we tested the accuracy on the training and test datasets and we see that it is 83.07% and 81.67%, respectively. The "out of sample" error is in agreement with the OOB estimate.

Next we display an error plot of the random forest model:
```{r Errorplot}
plot(ModRF)
```


```{R test}
print(ModRF)
```
The black solid line is the overall OOB error, the red dashed line is the prediction error for the first class/category (since we have a classification problem) of the variable "Class" - " <=50K", and the green dotted line represents the prediction error for the second category of"Class" - " >50K". From the error plot, we see that the errors do not decrease after the number of trees reaches a certain threshold, which is 100 trees. Hence, we try to build a RF model with 100 trees to check if we can achieve the same accuracy with a computationally less expensive model:


```{r using 100}
set.seed(1234)

start_time <- proc.time()
ModRF.small <- randomForest(formula = form, 
                            data = adult_train, 
                            ntree = 100)
time.rf.small <- proc.time() - start_time
time.rf.small

```
On our machine the computational time became faster.
```{r Print 100}
print(ModRF.small)
```

```{r}
mean(predict(ModRF.small, newdata = adult_train) == adult_train$Class)
mean(predict(ModRF.small, newdata = adult_test) == adult_test$Class)

```
From the code above,  the OOB error estimate increased insignificantly from 18.33% to 18.38% and the accuracy is a little smaller - 83.11% on the training data, and 81.59% on the test dataset. Next we show the confusion matrix:

```{r}
confusionMatrix(data = predict(ModRF.small, newdata = adult_test),
                reference = adult_test$Class)

```

```{r}
stat.accuracy.rf <- confusionMatrix(data = predict(ModRF.small, 
                                                   newdata = adult_test),
                                    reference = adult_test$Class)
```
