---
title: "FlowTaP Data Science Test"
author: "Uzo J Iwuagwu"
date: "May 24, 2017"
output: html_document
---

## 1. Engineering a Data Set
Create a list of numbers with the following properties: 
1) Minimum 100 distinct values, 
2) the Mean of all values is 1000 (+/- 0.5), 
3) the standard Deviation is 10 (+/- 0.1)

```{r  Engineering a Data Set}
set.seed(1)
value = rnorm(n = 500, mean = 1000, sd = 10)
value
round(mean(value), 1)
round(sd(value), 1)
```

## 2. Simulation

Imagine there is a country in which couples only want boys.
Couples continue to have children until they have their first boy. If they get a boy, they stop getting children.What is the long-term ratio of boys to girls in the country?

```{r Simulation}
p<-0
for (i in 1:10000){
  a<-0
  while(a != 1){   #Stops when having girls
    a<-as.numeric(rbinom(1, 1, 0.5))   #Simulation of a new birth with probability 0.5
    p=p+1   #Number of births
  }
}
(p-10000)/10000   #Ratio
```





